### 2012-02-26 paper.tex
Small typographic changes

CO oxidation is one of the most studied reactions due to its importance in both automotive catalysis and in the removal of trace CO in hydrogen streams. A large effort and a large number of experiments with a variety of surface sensitive techniques have been conducted on single crystal surfaces greatly improving the understanding of this reaction. Although single crystals surfaces are well suited model systems they can be very much different from industrial catalysts where supported nanoparticles are used to both increase surface area hence bringing material costs down and to increase the reactivity of the material by increasing the number of active sites.

Here we report on oscillatory behavior of CO oxidation on size-selected Pt nanoparticles supported on a SiO$_2$ substrate at atmospheric pressure. Oscillating reactions on single crystals at UHV conditions have been studied extensively and is know to originate from surface reconstructions of the surface during the oscillations\cite{Ertl2008}. It has previously been shown\cite{SALES1982} that CO oxidation at atmospheric pressures also shows oscillatory behavior on single crystals of Pt, Pd and Ir. Recent studies have proposed a mechanism on Pd \cite{Hendriksen2010} attributed to a cycle of surface oxide formation, oxide roughening and clean metal surface. CO oxidation oscillations has also been observed on Pt thin films\cite{Lund2000}, in macroscopic flow reactors\cite{Singh2010} and by wet chemical synthesis and subsequent annealing \cite{Carlsson2006}. Here we investigate CO oscillations in a microreactor model system on size selected Pt nanoparticles by high sensitivity mass spectrometry.

\section{Experimental Methods}
The experiments were performed in Si-based $\mu$-reactors\cite{Henriksen2009}. The reactor design makes sure that all gases exposed to the catalyst under investigation is measured by the QMA ensuring an extremely high sensitivity of the system.

@@

An example of a single very slow oscillation is seen in Figure~\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min.) where all CO is converted to CO$_2$ the reaction shows a fast deactivation. After the deactivation the sample quickly regains $\sim$30\% conversion. Hereafter the conversion increases slowly over time until approximately 50\% conversion is reached (at $\sim$1118\,min.). Shortly after reaching 50\% conversion the sample again ignites and convert all CO to CO$_2$. After a short full conversion period the sample again deactivates and the cycle repeats itself. The ratio between periods of high and low conversion are very different among samples, but for a given sample the ratio normally does not change even though the oscillation period consistently becomes longer as the experiment progresses.
\begin{figure*}
  \centering
  \includegraphics[width=17cm]{single_full_oscillation.png}
  \caption{A full oscillation period first showing a steep ignition of the sample followed by an almost immediate de-activation. For the next 65 minutes the sample slowly recovers activity until full conversion is reached again and the cycle repeats itself. This extremely long oscillation period is not commonly seen and was a result of carefull parameter tuning. Typical oscillation periods are normally between $\sim$30\,s and $\sim$30\,minutes.}
  \label{fgr:full_oscillation}
\end{figure*}

@@

Figure~\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations extracted from the same four day measurement. Data from shortly after the initial treatment, after 2800\,min. of oscillation time and after 5900\,min. are shown. After 5900\,min. the oscillations are more irregular which is consistent with the data shown in Figure \ref{fgr:long_measurement}. Also, oscillations in between full conversion oscillations with smaller amplitude and much higher frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}[h]
  \centering
  \includegraphics[width=9cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from the 4 day long experiment. The oscillation period become more irregular with time while the total integrated conversion remains constant. Furthermore, small oscillations in between full conversion cycles become more prominent as time progresses.}
  \label{fgr:extracts}
\end{figure}

The oscillation frequency becomes more irregular with progressing experiment time as shown in Figure \ref{fgr:extracts}. However, the ratio between CO and CO$_2$ integrated over an full period remains constant throughout the entire experiment time.

@@

\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{temperature_dependence.png}
  \caption{After about 10 hours of oscillations at 205$^\circ$C the temperature is first lowered by 5$^\circ$C and soon after increased by 2$^\circ$C. Before the temperature step the oscillation period was approximately 500\,s and after the step the period is approximately 5000\,s.}
  \label{fgr:temperature_dependence}
\end{figure}
A very strong temperature dependence on the oscillation frequency was also found as shown in Figure~\ref{fgr:temperature_dependence}. By changing the temperature 5$^\circ$C the oscillation period was changed by more than a factor of 10. The magnitude of the change is not completely consistent across all measured samples but all samples do show a very large temperature dependency where in some cases a provoked small change in temperature will consistently turn on and off the oscillations.

The pressure dependence of the oscillations were also investigated. In the pressure range of 0.1\,bar to 1\,bar no change of the oscillation frequency or qualitative behavior that could be attributed to the pressure was found. Similarly, only a weak dependence on the CO/O$_2$ ratio was observed. At a CO/O$_2$ ratio of $\sim0.025$ only very small amplitude oscillations were observed. Increasing the CO/O$_2$ showed clear oscillations with a significant increase in oscillation amplitude compared to lower CO concentrations. Generally, a tendency towards slower oscillation periods for higher CO/O$_2$ ratio was observed. Richer CO mixtures than 0.175 did not produce any oscillations in our system.

@@

It has previously been shown using STM \cite{Hendriksen2002}, FT-IR \cite{Carlsson2006}, Monte Carlo simulations \cite{Zhdanov2002} as well as DFT \cite{Gong2004} that the activity of Pt towards CO oxidation at atmospheric pressures is highly dependent on the oxidation state of the surface. A detailed proposal for the reaction mechanism on Pd has been developed by Hendriksen \textit{et. al.}\cite{Hendriksen2010} and our data is consistent with this model. We are currently investigating new methods for gaining insight into the nature of the nanoparticles in the reactor but the low coverage and the design of the reactor makes this a challenging task.

According to the proposed model in ref.~\citenum{Hendriksen2010} the bare metal surface is less active than the oxidized surface. Oscillations originate from a switch between an oxidized surface and the bare metal surface. Initially, with CO present in the inlet gas a smooth metal surface is exposed due to the general high adsorption energy of CO on metal surfaces. As the Pt nanoparticles convert the CO to CO$_2$ the partial pressure of CO will decrease and a Pt oxide will start to form in the high partial pressures of O$_2$. According to \cite{Hendriksen2010} a slow roughening of the particles will occur when the reaction occurs on Pt-oxide resulting in the formation of an increasingly rough oxide surface. As the oxide becomes rougher the bare metal surface is increasingly favored resulting in a return to the lower rate rough metal state. As the rate decreases the CO partial pressure increases which results in a transition from rough metal surface to a smooth metal surface. When the sample is sufficiently smooth it will again oxidize and thus complete the oscillation cycle.

@@

Nanoparticles of several difference sizes where tested which all showed oscillations. However, no change in duty cycle or frequency that could be attributed to the size of the nanoparticles was found. This may be due to the fact that even nominally identical samples also show a large variation in oscillation frequency and duty cycle thus possibly shadowing a nanoparticle size effect.

>>>>

CO oxidation is one of the most studied reactions due to its importance in both automotive catalysis and in the removal of trace CO in hydrogen streams. A large effort and a large number of experiments with a variety of surface sensitive techniques have been conducted on single crystal surfaces greatly improving the understanding of this reaction. Single crystals surfaces are well suited model systems due to the high degree of achievable control of parameters compared to industrial catalysts. Here we report on studies in a microreactor model system of Pt nanoparticles. The size-selected nanoparticles supported on a SiO$_2$ substrate show oscillatory behavior of CO oxidation at atmospheric pressure.

Oscillating reactions on single crystals at UHV conditions have been studied extensively and is know to originate from surface reconstructions of the surface during the oscillations\cite{Ertl2008}. It has previously been shown\cite{SALES1982} that CO oxidation at atmospheric pressures also shows oscillatory behavior on single crystals of Pt, Pd and Ir. Recent studies have proposed a mechanism on Pd \cite{Hendriksen2010} attributed to a cycle of surface oxide formation, oxide roughening and clean metal surface. CO oxidation oscillations has also been observed on Pt thin films\cite{Lund2000}, in macroscopic flow reactors\cite{Singh2010} and on Pt nanoparticles synthesized from wet chemical synthesis and subsequent annealing \cite{Carlsson2006}. Here we investigate CO oscillations in a microreactor model system on size selected Pt nanoparticles by high sensitivity mass spectrometry.

\section{Experimental Methods}
The experiments were performed in Si-based $\mu$-reactors\cite{Henriksen2009}. The reactor design makes sure that all gases exposed to the catalyst under investigation is measured by the quadropole mass spectrometer (QMA) ensuring an extremely high sensitivity of the system.

@@

An example of a single very slow oscillation is seen in Figure~\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min.) where all CO is converted to CO$_2$ the reaction shows a fast deactivation. After the deactivation the sample quickly regains $\sim$30\% conversion (at $\sim$1040\,min.). Hereafter the conversion increases slowly over time until approximately 50\% conversion is reached (at $\sim$1118\,min.). Shortly after reaching 50\% conversion the sample again ignites and convert all CO to CO$_2$. After a short full conversion period the sample again deactivates and the cycle repeats itself.
\begin{figure*}
  \centering
  \includegraphics[width=17cm]{single_full_oscillation.png}
  \caption{A full oscillation period first showing a steep ignition of the sample followed by an almost immediate de-activation. For the next 65 minutes the sample slowly recovers activity until full conversion is reached again and the cycle repeats itself. This extremely long oscillation period is not commonly seen and was a result of carefull parameter tuning. Typical oscillation periods are normally between $\sim$30\,s and $\sim$30\,minutes.}
  \label{fgr:full_oscillation}
\end{figure*}

@@

Figure~\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations extracted from the same four day measurement. Data from shortly after the initial treatment, after 2800\,min. of oscillation time and after 5900\,min. are shown. After 5900\,min. the oscillations are more irregular which is consistent with the data shown in Figure \ref{fgr:long_measurement}. Also, oscillations in between full conversion oscillations with smaller amplitude and much higher frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}[h]
  \centering
  \includegraphics[width=9cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from the 4 day long experiment. The oscillation period becomes more irregular with time while the total integrated conversion remains constant. Furthermore, small oscillations in between full conversion cycles become more prominent as time progresses.}
  \label{fgr:extracts}
\end{figure}

The oscillation frequency becomes more irregular with progressing experiment time as shown in Figure \ref{fgr:extracts}. However, the ratio between CO and CO$_2$ integrated over an full period remains constant throughout the entire experiment time, i.e. the samples produces the same amount of CO$_2$ independent of oscillation frequency.

@@

\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{temperature_dependence.png}
  \caption{After about 10 hours of oscillations at 205$^\circ$C the temperature is first lowered by 5$^\circ$C and soon after increased by 2$^\circ$C. Before the temperature step the oscillation period was approximately 500\,s and after the step the period is approximately 5000\,s.}
  \label{fgr:temperature_dependence}
\end{figure}
A very strong temperature dependence on the oscillation frequency was also found as shown in Figure~\ref{fgr:temperature_dependence}. By changing the temperature 5$^\circ$C the oscillation period was changed by more than a factor of 10. The magnitude of the change is not completely consistent across all measured samples but all samples show a very large temperature dependency where in some cases a provoked small change in temperature will consistently turn on and off the oscillations.

The pressure dependence of the oscillations was also investigated. In the pressure range of 0.1\,bar to 1\,bar no change of the oscillation frequency or qualitative behavior that could be attributed to the pressure was found. Similarly, only a weak dependence on the CO/O$_2$ ratio was observed. At a CO/O$_2$ ratio of $\sim0.025$ only very small amplitude oscillations was observed. Increasing the CO/O$_2$ showed clear oscillations with a significant increase in oscillation amplitude compared to lower CO concentrations. Generally, a tendency towards slower oscillation periods for higher CO/O$_2$ ratio was observed. Richer CO mixtures than 0.175 did not produce any oscillations in our system.

@@

It has previously been shown using STM \cite{Hendriksen2002}, FT-IR \cite{Carlsson2006}, Monte Carlo simulations \cite{Zhdanov2002} as well as DFT \cite{Gong2004} that the activity of Pt towards CO oxidation at atmospheric pressures is highly dependent on the oxidation state of the surface. We are currently investigating new methods for gaining insight into the nature of the nanoparticles in the reactor but the low coverage and the design of the reactor makes this a challenging task. A detailed proposal for the reaction mechanism on Pd has been developed by Hendriksen \textit{et. al.}\cite{Hendriksen2010} and our data is consistent with this model.

According to the proposed model in ref.~\citenum{Hendriksen2010} the bare metal surface is less active than the oxidized surface. Oscillations originate from a switch between an oxidized surface and the bare metal surface. Initially, with CO present in the inlet gas a smooth metal surface is exposed due to the general high adsorption energy of CO on metal surfaces. As the Pt nanoparticles convert the CO to CO$_2$ the partial pressure of CO will decrease and a Pt oxide will start to form in the high partial pressures of O$_2$. After the oxidation of the nanoparticles they will roughen slowly during CO oxidation resulting in the formation of an increasingly rough oxide surface. As the oxide becomes rougher the bare metal surface is increasingly favored resulting in a return to the lower rate rough metal state. As the rate decreases the CO partial pressure increases which results in a transition from rough metal surface to a smooth metal surface. When the sample is sufficiently smooth it will again oxidize and thus complete the oscillation cycle.

@@

Nanoparticles of several different sizes where tested which all showed oscillations. However, no change in duty cycle or frequency that could be attributed to the size of the nanoparticles was found. This may be due to the fact that even nominally identical samples also show a large variation in oscillation frequency and duty cycle thus possibly shadowing a nanoparticle size effect.
